---
title: "tCPOE Calculations"
author: "Michael Egle"
date: "10/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(modelr)
```


First load in the metadata
```{r}
playinfo <- read_csv("plays.csv")
gameinfo <- read_csv("games.csv")
playerinfo <- read_csv("players.csv")
receivers <- read_csv("targetedReciever.csv")
```

These event types will be useful in determining the ball's position at important points during the play
```{r}
pass_arrival_events <- c("pass_outcome_incomplete",
                         "pass_arrived",
                         "pass_outcome_intercepted",
                         "pass_outcome_caught",
                         "pass_outcome_touchdown")
```

```{r}
pass_release_events <- c("pass_forward",
                         "pass_shovel")
```

Flatten the data a little bit. We want just one observation per play to train our model
```{r}
train <- data.frame()

for (i in 1:17)
{
  file_name <- paste0("week", i, ".csv")
  
  week <- read_csv(file_name)
  
  
  week <- week %>%
    distinct()
  
  if(i == 3)
  {
    week <- week %>%
      filter(!(gameId == 2018092304 & playId == 1687))
  }
  
  if(i == 6)
  {
    week <- week %>%
      filter(!(gameId == 2018101410 & playId == 431))
  }
  
  if(i == 9)
  {
    week <- week %>%
      filter(!(gameId == 2018110410 & playId == 3640))
  }
  
  if(i == 10)
  {
    week <- week %>%
      filter(!(gameId == 2018111106 & playId == 2650))
  }
  
  week <- week %>%
    inner_join(playinfo) %>%
    inner_join(gameinfo) %>%
    inner_join(receivers) %>%
    mutate(team = ifelse(team == "home", homeTeamAbbr, visitorTeamAbbr))
  
  week <- week %>%
    group_by(gameId, playId) %>%
    mutate(spike = sum(event == "qb_spike") > 0,
           off_players = sum(team == possessionTeam & displayName != "Football"),
           def_players = sum(team != possessionTeam & displayName != "Football")) %>%
    filter(!spike) %>%
    select(-spike) %>%
    ungroup() %>%
    filter(off_players > 0 & def_players > 0)
  
  week <- week %>%
    group_by(gameId, playId) %>%
    mutate(ball_in_play_release = sum(displayName == "Football" & event %in% pass_release_events) > 0) %>%
    filter(ball_in_play_release) %>%
    mutate(x_ball_start = x[displayName == "Football" & event %in% pass_release_events],
           y_ball_start = y[displayName == "Football" & event %in% pass_release_events],
           ball_in_play_arrival = sum(displayName == "Football" & event %in% pass_arrival_events) > 0) %>%
    filter(ball_in_play_arrival) %>%
    mutate(arrival_frame = max(frameId[event %in% pass_arrival_events]),
           x_ball_end = x[displayName == "Football" & frameId == arrival_frame],
           y_ball_end = y[displayName == "Football" & frameId == arrival_frame],
           geom_air_yards = sqrt((x_ball_end - x_ball_start)^2 + (y_ball_end - y_ball_start)^2),
           max_ball_speed = max(s[displayName == "Football"]))
  
  week_arrive <- week %>%
    filter(frameId == arrival_frame) %>%
    mutate(is_intended_receiver = nflId == targetNflId,
           is_intended_receiver = ifelse(is.na(is_intended_receiver), F, is_intended_receiver))
  
  week_arrive <- week_arrive %>%
    filter(!is.na(targetNflId)) %>%
    group_by(gameId, playId) %>%
    mutate(found_receiver = sum(is_intended_receiver) > 0) %>%
    filter(found_receiver) %>%
    mutate(wr_x = x[is_intended_receiver],
           wr_y = y[is_intended_receiver],
           distance_to_wr = sqrt((wr_x - x)^2 + (wr_y - y)^2),
           ball_wr_distance = sqrt((wr_x - x_ball_end)^2 + (wr_y - y_ball_end)^2),
           distance_to_ball = sqrt((x_ball_end - x)^2 + (y_ball_end - y)^2),
           bound_wr_distance = min(wr_x, 120 - wr_x, wr_y, 53.333 - wr_y)) %>%
    filter(def_players > 0 & off_players > 0) %>% # some plays don't have any off or def players
    mutate(min_def_distance = min(distance_to_wr[team != possessionTeam & displayName != "Football"]),
           nearest_def = displayName[distance_to_wr == min_def_distance]) %>%
    filter(displayName == nearest_def) %>%
    mutate(complete = ifelse(passResult == "C", 1, 0)) %>%
    mutate(ab.x = x - x_ball_end,
         ab.y = y - y_ball_end,
         bc.x = wr_x - x,
         bc.y = wr_y - y,
         ab = sqrt(ab.x^2 + ab.y^2),
         bc = sqrt(bc.x^2 + bc.y^2),
         theta = acos(((ab.x * bc.x) + (ab.y * bc.y)) / (ab * bc)),
         theta = ifelse(theta > pi, theta - pi, theta)) %>% # Have the angles go from [0, pi)
    select(gameId, playId, nflId, complete, nearest_def, distance_to_wr, distance_to_ball,
           ball_wr_distance, geom_air_yards, max_ball_speed, team, numberOfPassRushers, theta,
           epa, bound_wr_distance, targetNflId)
  
  train <- train %>%
    bind_rows(week_arrive)
  
  print(paste("Week", i, "completed."))
}
```

```{r}
cpoe_model <- glm(complete ~ distance_to_wr + theta + distance_to_ball + ball_wr_distance +
                    geom_air_yards + max_ball_speed + bound_wr_distance,
                  data = train,
                  family = "binomial")
```

```{r}
summary(cpoe_model)
```

Train the model
```{r}
train <- train %>%
  add_predictions(cpoe_model, var = "xcomp", type = "response") %>%
  mutate(cpoe = complete - xcomp)
```

Aggregate the train dataframe down to the player level. Feel free to mess around with the parameters of this chunk as you wish. I'd recommend filtering to a minimum of 20 targets, however this line is not currently included in the code.
```{r}
train %>%
  filter(distance_to_wr < 5) %>%
  inner_join(playinfo) %>%
  filter(!is.na(cpoe)) %>%
  filter(cpoe < 0.99) %>%
  group_by(nflId) %>%
  summarize(name = head(nearest_def, 1),
            team = tail(team, 1),
            tCPOE = mean(cpoe),
            targets = n(),
            catches = sum(complete),
            EPA = mean(epa)) %>%
  arrange(tCPOE) %>%
  ungroup() -> secondary
```

Export to csv
```{r}
write_csv(secondary, "secondary.csv")
```
